---
title: "LifeGenMon Fagus SNP analysis - Data filtering options"
author: "Nikos Tourvas"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
bibliography: library.bib
output:
  html_notebook:
    code_folding: hide
    theme: united
    toc: yes
    toc_float: yes
    toc_depth: 3
  html_document:
    code_folding: hide
    theme: united
    toc: yes
    ## toc_float: yes
  pdf_document:
    toc: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
```

```{r}
source("import_snp.R")
source("barplot_missing.R")
source("barplot_N_alleles.R")
```


# Case 1

```{r}
threshold_loci <- 0.1
threshold_ind <- 0.1
maf <- 0.05

obj <- import.snp.fagus(threshold_loci = threshold_loci,
                        threshold_ind = threshold_ind,
                        maf = maf)
```

### Data filtering thresholds
```{r include=T}
paste("SNP loci:", threshold_loci*100, 
      "% missing data across populations", sep=" ")
yintercept = 20

paste("Individuals:", threshold_ind*100, "% missing data", sep=" ")


paste("SNP loci with minor allele frequency <", maf)
```

```{r include=T}
### Data set details after data filtering:

paste(nInd(obj), "individuals", "and", nLoc(obj), "loci", sep=" ")
```

```{r, include=T}
barplot.missing(obj)
```

```{r cache=T, dpi=300, fig.width=5, fig.height=3, include=T}
a1 <- barplot.N.alleles(obj)
a1$plot
```

### Private alleles per cohort - level Country/Pop
```{r, include=T}
priv <- poppr::private_alleles(obj, report = "data.frame")

if(is.data.frame(priv)){
  
ggplot(priv) + geom_tile(aes(x = population,
                                y = allele,
                                fill = count)) +
  ggtitle("Private alleles per population") +
  scale_fill_viridis_c()
  
}
```
### Private alleles per cohort - level Country
```{r, include=T}
setPop(obj) <- ~Country

priv <- poppr::private_alleles(obj, 
                                  report = "data.frame")


if(is.data.frame(priv)){

ggplot(priv) + geom_tile(aes(x = population,
                                y = allele,
                                fill = count)) +
  ggtitle("Private alleles per cohort") +
  scale_fill_viridis_c()
  
}

setPop(obj) <- ~Country/Pop
```

```{r}
setPop(obj) <- ~Country

priv_count1 <- poppr::private_alleles(obj, 
                                  report = "data.frame",
                                  count.alleles = F)
priv_count1$maf <- maf

# ggplot(priv_count, aes(x=population, y=count)) +
#   geom_bar(stat="identity")

setPop(obj) <- ~Country/Pop
```


# Case 2 - No MAF filter

```{r}
threshold_loci <- 0.1
threshold_ind <- 0.1
maf <- 0.03

obj <- import.snp.fagus(threshold_loci = threshold_loci,
                        threshold_ind = threshold_ind,
                        maf = maf)
```

### Data filtering thresholds
```{r include=T}
paste("SNP loci:", threshold_loci*100, 
      "% missing data across populations", sep=" ")


paste("Individuals:", threshold_ind*100, "% missing data", sep=" ")


paste("SNP loci with minor allele frequency <", maf)
```

```{r include=T}
### Data set details after data filtering:

paste(nInd(obj), "individuals", "and", nLoc(obj), "loci", sep=" ")
```

```{r, include=T}
barplot.missing(obj)
```

```{r cache=T, dpi=300, fig.width=5, fig.height=3, include=T}
a2 <- barplot.N.alleles(obj)
a2$plot
```

### Private alleles per cohort - level Country/Pop
```{r, include=T}
priv <- poppr::private_alleles(obj, report = "data.frame")

if(is.data.frame(priv)){
  
ggplot(priv) + geom_tile(aes(x = population,
                                y = allele,
                                fill = count)) +
  ggtitle("Private alleles per population") +
  scale_fill_viridis_c()
  
}
```

### Private alleles per cohort - level Country
```{r, include=T}
setPop(obj) <- ~Country

priv <- poppr::private_alleles(obj, 
                                  report = "data.frame")


if(is.data.frame(priv)){

ggplot(priv) + geom_tile(aes(x = population,
                                y = allele,
                                fill = count)) +
  ggtitle("Private alleles per cohort") +
  scale_fill_viridis_c()
  
}

setPop(obj) <- ~Country/Pop
```

```{r}
setPop(obj) <- ~Country

priv_count2 <- poppr::private_alleles(obj, 
                                  report = "data.frame",
                                  count.alleles = F)
priv_count2$maf <- maf

# ggplot(priv_count, aes(x=population, y=count)) +
#   geom_bar(stat="identity")

setPop(obj) <- ~Country/Pop
```

# Case 3

```{r}
threshold_loci <- 0.1
threshold_ind <- 0.1
maf <- 0.01

obj <- import.snp.fagus(threshold_loci = threshold_loci,
                        threshold_ind = threshold_ind,
                        maf = maf)
```

### Data filtering thresholds
```{r include=T}
paste("SNP loci:", threshold_loci*100, 
      "% missing data across populations", sep=" ")


paste("Individuals:", threshold_ind*100, "% missing data", sep=" ")


paste("SNP loci with minor allele frequency <", maf)
```

```{r include=T}
### Data set details after data filtering:

paste(nInd(obj), "individuals", "and", nLoc(obj), "loci", sep=" ")
```

```{r, include=T}
barplot.missing(obj)
```

```{r cache=T, dpi=300, fig.width=5, fig.height=3, include=T}
a3 <- barplot.N.alleles(obj)
a3$plot
```

### Private alleles per cohort - level Country/Pop
```{r, include=T}
priv <- poppr::private_alleles(obj, report = "data.frame")

if(is.data.frame(priv)){
  
ggplot(priv) + geom_tile(aes(x = population,
                                y = allele,
                                fill = count)) +
  ggtitle("Private alleles per population") +
  scale_fill_viridis_c()
  
}
```

### Private alleles per cohort - level Country
```{r, include=T}
setPop(obj) <- ~Country

priv <- poppr::private_alleles(obj, 
                                  report = "data.frame")


if(is.data.frame(priv)){

ggplot(priv) + geom_tile(aes(x = population,
                                y = allele,
                                fill = count)) +
  ggtitle("Private alleles per cohort") +
  scale_fill_viridis_c()
  
}

setPop(obj) <- ~Country/Pop
```

```{r}
setPop(obj) <- ~Country

priv_count3 <- poppr::private_alleles(obj, 
                                  report = "data.frame",
                                  count.alleles = F)
priv_count3$maf <- maf

# ggplot(priv_count, aes(x=population, y=count)) +
#   geom_bar(stat="identity")

setPop(obj) <- ~Country/Pop
```

# Case 4

```{r}
threshold_loci <- 0.1
threshold_ind <- 0.1
maf <- 0

obj <- import.snp.fagus(threshold_loci = threshold_loci,
                        threshold_ind = threshold_ind,
                        maf = maf)
```

### Data filtering thresholds
```{r include=T}
paste("SNP loci:", threshold_loci*100, 
      "% missing data across populations", sep=" ")


paste("Individuals:", threshold_ind*100, "% missing data", sep=" ")


paste("SNP loci with minor allele frequency <", maf)
```

```{r include=T}
### Data set details after data filtering:

paste(nInd(obj), "individuals", "and", nLoc(obj), "loci", sep=" ")
```

```{r, include=T}
barplot.missing(obj)
```

```{r cache=T, dpi=300, fig.width=5, fig.height=3, include=T}
a4 <- barplot.N.alleles(obj)
a4$plot
```

### Private alleles per cohort - level Country/Pop
```{r, include=T}
priv <- poppr::private_alleles(obj, report = "data.frame")

if(is.data.frame(priv)){
  
ggplot(priv) + geom_tile(aes(x = population,
                                y = allele,
                                fill = count)) +
  ggtitle("Private alleles per population") +
  scale_fill_viridis_c()
  
}
```

### Private alleles per cohort - level Country
```{r, include=T}
setPop(obj) <- ~Country

priv <- poppr::private_alleles(obj, 
                                  report = "data.frame")


if(is.data.frame(priv)){

ggplot(priv) + geom_tile(aes(x = population,
                                y = allele,
                                fill = count)) +
  ggtitle("Private alleles per cohort") +
  scale_fill_viridis_c()
  
}

setPop(obj) <- ~Country/Pop
```

```{r}
setPop(obj) <- ~Country

priv_count4 <- poppr::private_alleles(obj, 
                                  report = "data.frame",
                                  count.alleles = F)
priv_count4$maf <- maf

# ggplot(priv_count, aes(x=population, y=count)) +
#   geom_bar(stat="identity")

setPop(obj) <- ~Country/Pop
```



# Overall

```{r}
priv_count1 <- as.data.frame(priv_count1)
# priv_count1 <- priv_count1 %>% 
#   add_column("population") %>%
#   add_column("count") %>% 
#   add_column("allele")
# priv_count1 <- priv_count1 %>%
#   add_row(population = "DE", count = 0, allele=0) %>%
#   add_row(population = "GR", count = 0, allele=0) %>%
#   add_row(population = "SI", count = 0, allele=0)
priv_count1$population <- NA
priv_count1$allele <- NA
priv_count1$count <-0

priv_count2 <- as.data.frame(priv_count2)
# priv_count2 <- priv_count2 %>% 
#   add_column("population") %>% 
#   add_column("count") %>% 
#   add_column("allele")
# priv_count2 <- priv_count2 %>% 
#   add_row(population = "DE", count = 0, allele=0) %>% 
#   add_row(population = "GR", count = 0, allele=0) %>% 
#   add_row(population = "SI", count = 0, allele=0)

priv_count2$population <- NA
priv_count2$allele <- NA
priv_count2$count <-0
```


```{r, include=T, fig.width=5, fig.height=2.5}
priv_count <- rbind(priv_count1,
                    priv_count2,
                    priv_count3,
                    priv_count4)

# ggplot(priv_count, aes(x=population, y=count)) +
#   geom_bar(stat="identity") +
#   facet_wrap(~maf)

priv_count <- priv_count %>% 
  filter(count > 0) %>% 
  group_by(maf) %>% 
  summarise(count = n())

priv_count <- priv_count %>% 
  add_row(maf = 0.03, count = 0) %>% 
  add_row(maf = 0.05, count = 0)

ggplot(priv_count, aes(x=as.factor(maf), y=count)) +
  geom_bar(stat="identity",
           # color = "black", fill = "white",
                 width = .5) +
        ggtitle("Total Number of Private Alleles per MAF Threshold") +
  theme_classic() +
  labs(x = "MAF") + 
  theme(axis.text.x = element_text(size = 14), 
    axis.text.y = element_text(size = 14)) +
  scale_y_continuous(breaks = seq(0, 8, 2))
  


ggplot(priv_count, aes(x=as.factor(maf), y=count, group=1)) +
  geom_point()+ 
  geom_line(color="black")
```

```{r fig.width=5, fig.height=2.5}
a <- rbind(a1[["data"]], a2[["data"]], a3[["data"]], a4[["data"]])

p1 <- ggplot(a, aes(x=as.factor(maf), y=count)) +
      geom_bar(stat="identity",
               # color = "black", fill = "white",
                     width = .5) +
            ggtitle("Total Number of Alleles per MAF Threshold") +
      theme_classic()

p1 +labs(x = "MAF") + 
    theme(axis.text.x = element_text(size = 14), 
    axis.text.y = element_text(size = 14))

p <- ggplot(a, aes(x=population, y=count)) +
  geom_bar(stat="identity") +
  facet_wrap(~maf, ncol = 4)

p + theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    axis.text.x = element_text(angle = 90))
```


In the previous 2 cases the missing data filter for SNP loci is applied on the whole data set (not for each population). This means that a 10% threshold will result in one or more populations ending up with higher percentage of missing data than 10% for certain loci.

For the following 2 cases the missing data filter for SNP loci is applied for each population.

# Case 3

```{r}
threshold_loci <- 0.1
threshold_ind <- 0.1
maf <- 0.05

obj <- import.snp.fagus2(threshold_loci = threshold_loci,
                        threshold_ind = threshold_ind,
                        maf = maf)
```

### Data filtering thresholds
```{r include=T}
paste("SNP loci:", threshold_loci*100, 
      "% missing data across populations", sep=" ")


paste("Individuals:", threshold_ind*100, "% missing data", sep=" ")


paste("SNP loci with minor allele frequency <", maf)
```

```{r include=T}
### Data set details after data filtering:

paste(nInd(obj), "individuals", "and", nLoc(obj), "loci", sep=" ")
```

```{r, include=T}
barplot.missing(obj)
```

```{r cache=T, dpi=300, fig.width=5, fig.height=3, include=T}
a3 <- barplot.N.alleles(obj)
a3$plot
```

# Case 4 - No MAF filter

```{r}
threshold_loci <- 0.1
threshold_ind <- 0.1
maf <- 0

obj <- import.snp.fagus2(threshold_loci = threshold_loci,
                        threshold_ind = threshold_ind,
                        maf = maf)
```

### Data filtering thresholds
```{r include=T}
paste("SNP loci:", threshold_loci*100, 
      "% missing data across populations", sep=" ")


paste("Individuals:", threshold_ind*100, "% missing data", sep=" ")


paste("SNP loci with minor allele frequency <", maf)
```

```{r include=T}
### Data set details after data filtering:

paste(nInd(obj), "individuals", "and", nLoc(obj), "loci", sep=" ")
```

```{r, include=T}
barplot.missing(obj)
```

```{r cache=T, dpi=300, fig.width=5, fig.height=3, include=T}
a4 <- barplot.N.alleles(obj)
a4$plot
```
